The Heavy Photon Search Test Detector

### 2013-11-03 hps-testrun-nim.tex
Added Stepans edits of ecal and daq. Updated turn-on and eoverp from Sho with new root style. Removed gain plot. Removed turn-on vs energy since I didnt have that plot with new style. Added references for clas IC and FADC

\subsection{SVT Digitization, Event Building and Data Transmission}
\label{sec:svtdaq}
The SVT data acquisition (DAQ) based on the Reconfigurable Cluster Element (RCE) and Cluster 
Interconnect (CI) concept developed at SLAC as generic building blocks for DAQ systems in high energy 
physics {\bf [need reference]}. 
The RCE is a generic computational building block, housed on a separate daughter card called 
Data Processing Module (DPM), that are realized on an ATCA front board called the Cluster On Board (COB), 
see Fig.~\ref{fig:cob}.
 \begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/svt_daq_module}
	\caption{Picture of the SVT DAQ COB board with four data processing daughter cards visible.}
	\label{fig:cob}
}
\end{center}
\end{figure}
The Gen-I RCE used in the Test run consisted of a Virtex~5 FPGA with 1~GB of DDR3 RAM. 
The outputs of 12 SVT half-modules 
are digitized on the Rear-Transition-Module (RTM), a custom board on the back side of the ATCA crate, 
interfacing the HPS readout to the generic DAQ components on the COB. 
A pre-amplifier converts the APV25 differential current output to a different voltage output scaled to the sensitive 
range of a 14-bit ADC operating at the system clock of 41.667~MHz. 
The RTM is organized into four sections with each section supporting three SVT half-module hybrids (15 
APV25 channels). The RTM also includes a 4-channel Fiber Optic module and supporting logic which can be 
used to interface to the JLab trigger system supervisor.
Each section of the RTM, up to three hybrids, is input to a DPM which apply thresholds for data reduction and 
organizes the sample data into UDP datagrams. The DPM also hosts an I2C controller used to configure and 
monitor the APV25 chips on those three hybrids. A single ATCA crate with two COB cards was used, one 
supporting four DPMs and one supporting 3 DPMs and one DPM that is configured as the trigger and data 
transmission module. The two COB cards and their DPMs are interconnected with a 10Gb/s switch 
card~\cite{Larsen:2011zb} which 
also hosts two 1Gb/s Ethernet interfaces to the external SVT DAQ PC.

@@

After a trigger is received, six samples of the corresponding output of the APV25 shaper circuit are digitized. 
The samples from every channel on a sensor surviving the 
data reduction algorithm (see Sec.~\ref{sec:svtdaq}) are the basis for offline hit reconstruction. 
The six samples of the APV25 shaper output 
are fitted to an ideal $CR-RC$ function to extract the amplitude and the $t_0$ of the hit. 
The typical pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape} also demonstrates that the SVT 
was well timed in to the trigger with the rise of the pulse at the 3rd sampling point.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/08062012_run1351_samples_vs_amplitude}
    	\caption{The six pedestal subtracted samples from individual channels associated with a hit on a 
	track.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}

@@

\section{Electromagnetic Calorimeter}
\label{ecal}

The electromagnetic calorimeter just downstream of the tracker uses detectors with comparably short 
live time and high rate capability. It performs two essential functions for the experiment: triggering and 
electron identification. The device is highly segmented. It is fast, able to readout at rates comparable to 
those in the tracker, and able to provide good spatial and energy information to the trigger electronics. Like 
the tracker system, the electromagnetic calorimeter is split to avoid impinging on the “dead zone”. The beam 
and radiative secondaries pass through the calorimeter in vacuum, to avoid generating unnecessary backgrounds.

@@

\subsection{Layout}
The modules in the ECal are arranged in two formations, as shown in Fig.~\ref{fig:ecal}. 
\begin{center}
{\small
\begin{figure*}[]
\includegraphics[width=7cm]{figures/ECal}
\includegraphics[width=7cm]{figures/ecal_assembly}
\caption{Schematic layout view of the two halves of the ECal (left) and a picture of one half being assembled before installation on beam line (right). }
\label{fig:ecal}
\end{figure*}
}
\end{center}
There are 5 layers in each formation. Each layer has $46$ crystals except for the layers closest to the beam 
plane where 9 modules were removed to allow a larger opening for the outgoing electron and photon beams. 
The ECal is mounted downstream of the analyzing dipole magnet at the distance of about $137$~cm from the 
upstream edge of the magnet. The two ECal modules are positioned just above and below the ECal vacuum 
chamber, through which the beam, radiated photons, and the wall of flame will pass
unimpeded. The innermost edge of the crystals is just $2$~cm from the beam.

@@

\subsubsection{PbWO$_4$ Crystals}
The ECal PbWO$_{4}$ crystals, from the same crystals that were used in the IC. The IC calorimeter was assembled from 424 lead tungstate tapered crystals with dimensions of $13.3\times13.3$~mm$^2$ (front face), $16\times16$~mm$^2$ (rear face) and 160~mm length. The crystals were fabricated with very tight tolerances: $19~\mu$m for 13.3~mm dimension and $13~\mu$m for the 16~mm dimension. Each crystal is wrapped in VM2000 multilayer polymer mirror film.

\subsubsection{Support Structure}
In order to stabilize the 
calorimeter's performance, the crystals, APDs, and amplifiers are enclosed within a temperature controlled 
environment, held constant at the level of $1^\circ~$F.

\subsubsection{Avalanche Photodiode Readout and Preamplifiers}
\label{sec:apd}
Avalanche Photodiodes (APD) S8664-55 produced by Hamamatsu Corp. are used as photodetectors. 
They have a $5\times5$~mm$^2$ active area and quantum efficiency of about 75\%. Single APDs were 
centered on the back ends of the crystals using MeltMount 1.7 thermal plastic adhesive. 
The low? amplitude signals of APDs are amplified using custom?made preamplifier boards. 
Figure~\ref{fig:ecal-module} shows a schematic view of the module assembly.
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}
The gain uniformity of the APDs is on the order of XXXX\%; with APD high voltage bias varying by as 
much as 100V. Since only a limited number of high voltage channels were available during the test, 10 
channels were grouped together with an attempt to group modules with similar gains.

The signals from the preamplifier boards were sent, through a $2:1$ signal splitter to a time-to-digital 
converter and the Flash ADC readout boards discussed in Sec.~\ref{sec:fadc}.

\subsection{Calibration and Performance}
\label{sec:ecal_calibration}
Of 442 modules, 39 were disabled or disconnected and were not read out by the DAQ. 
13 of these were not read out because of a shortage of FADC readout boards (see Sec.~\ref{sec:fadc}).
The remainder either had no HV bias on the APD, or were disabled in the FADC software due to noise.
In the data, we identified two types of abnormal channels. 
One FADC was not sending trigger signals correctly, resulting in low efficiency.
 This affected the 13 channels read out by that FADC. 5 channels were diagnosed as noisy because they 
 had a high incidence of hits out of coincidence with the trigger. A large number of channels were originally 
 misidentified as noisy because they had much higher hit occupancy than neighboring channels.
Gain calibration shows that these channels have high gain (and thus lower energy threshold) but are otherwise normal. The abnormal channels were ignored in analysis in order to simplify comparison with Monte Carlo. This leaves 385 useful channels---87\% of the ECal.

@@

We calibrate gain of the individual ECal channels using the SVT measurement of track momentum and 
comparison to Monte Carlo simulation.  We disable all SVT and ECal channels in the simulation that were 
inoperable or noisy in the test run, so any efficiency or bias effects that affect the real data should be reflected 
in the simulation as well; then we use a formula to compute the ``weighted E/p'' for a crystal, representing the average E/p for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$
, and iteratively adjust the gains until the weighted E/p is equal for test run data and simulation.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/ecalgainplots_corr_sim}
	\includegraphics[width=7cm]{figures/gains}
	\caption{Weighted E/p from Monte Carlo simulation (top), calibrated values of gain in units of MeV 
	per ADC count (bottom).}
	\label{fig:gains}
}
\end{center}
\end{figure}

@@

\subsection{Monitoring}

@@

\section{Trigger and Data Acquisition}
The HPS Test apparatus DAQ handles the acquisition of data for two sub-detectors: the 
SVT and ECal . It employs two DAQ architectures: the SVT is readout with Advanced 
Telecom Communications Architecture (ATCA) hardware while the ECal and Muon System use VXS based 
hardware. The trigger system receives input from the ECal and distributes a trigger signal 
to all detector sub- systems to read out a selected event.

\subsection{Electronics}
\label{sec:electronics}

\subsubsection{Flash ADC Readout Board}
\label{sec:fadc}
The analog signals from the individual APD's of the ECal (shaped and amplified as described in 
Sec.~\ref{sec:aid}) are input to a single channel on the 16-channel JLab FADC250 VXS module (FADC), 
shown in Fig. ~\ref{fig:fadc}. 
\begin{figure}[t]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/FADC250_Photo_001.jpg}
\caption{A Jefferson Lab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
Three 20-slot VXS crates are needed to accommodate the system: one for each half of the ECal with 221 
channels and one for the Muon System with a total of 232 channels.

The FADCs store 12-bit digitized samples at 250~MHz in 8~$\mu$s deep pipelines. 
When a trigger is received, the appropriate part of the pipeline is accessed. If a FADC   
signal exceeds a predefined threshold within that time window, the integrated amplitude of a pre-defined 
number of samples before (NSB) and after (NSA) the signal passed threshold, in addition to the time, are 
recorded. 
%as explained in Fig.~\ref{fig:hps_trigger_data}. 
This scheme significantly compresses the data input 
to the FADC. During data analysis, a pedestal value is subtracted to obtain the actual summed energy.
The main characteristics of the FADC are:
\begin{itemize}
\item 12-bit digitizer with sampling rate of 250~Msps, 
\item 50$\Omega$ termination input, 
\item front-end input range:  -0.5V, -1V or -2V (sufficient to avoid signal clipping for large pulse heights),
\item nominal charge resolution between 10-39~fC per ADC (see Tab.~\ref{tab:charge_resolution}).
\end{itemize}
\begin{center}
{\small
\begin{table}[h]
\centering
\begin{tabular}{|c|c|}
\hline
Input range & Nominal charge resolution\\
(V) & (fC per ADC count)\\\hline
-0.5 & 9.76  \\\hline
-1.0 & 19.53  \\\hline
-2.0 & 39.06 \\\hline
\end{tabular}
\caption{Nominal FADC charge resolution for different front-end input ranges.}
\label{tab:charge_resolution}
\end{table}
}
\end{center}
As shown in Fig.~\ref{fig:hps_trigger_data}, the FADC has two parallel data paths: the readout and trigger paths. The trigger path runs continuously to report hits to the trigger system. The readout path only reports hits to the DAQ when the FADC receives a trigger.

For the readout path, every FADC has the following parameters:
 \begin{itemize}
 \item the number of samples integrated before the signal crossed threshold ($NSB$), 
 \item the number of samples integrated after the signal crossed threshold ($NSA$),
 \item the readout threshold, measured in ADC counts.
 \end{itemize}
The number of samples for a given channel integration is the sum of $NSB+NSA$ samples. It is a fixed gate width pulse integration with no pedestal subtraction where the sum is stored in a 17-bit register for readout (pedestal subtraction happens offline). 
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/hps_trigger_data}
\caption{\small{FADC data paths}}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}
For the trigger path, every channel has, in addition to $NSB$ and $NSA$: 
\begin{itemize}
% \item number of samples integrated before the threshold crossing (NSB),
 %\item number of samples integrated after the  threshold crossing (NSA),
 \item trigger threshold, measured in ADC counts, 
 \item a pedestal, 
 \item a conversion factor (gain) that converts the ADC counts to energy in MeV (with 13 bits:  from 0 to 8191~MeV), 
 \item an energy discriminator threshold (minimum energy cutoff).
 \end{itemize}
Note that the threshold for the trigger path can be set independently from the readout threshold.
The pedestal value is subtracted from the integrated sum over $NSB+NSA$ samples and 
converted to MeV units using a supplied gain conversion factor. The energy discriminator can be used to cut 
off low energy pulses before reporting to the Crate Trigger Processor. 
The values reported to the Crate Trigger Processor are the 13-bit pulse energy and the time at which the 
pulse crossed the threshold. Data for every channel is sent every 32~ns (if there is no hit a 0 energy pulse is 
sent) which sets a worst case double pulse resolution of 32~ns for individual channels, but less if pulses 
occur in adjacent 32~ns windows.

\subsubsection{Trigger Boards}
Describe CTP and SSP?

\subsection{Trigger}
\label{sec:trigger}
The trigger system is designed to efficiently select \ee{} events by 
using information from the ECal . The trigger looks for time coincidences of clusters in the top and bottom half 
of the ECal.  The trigger system can be broken down into three sections (see Fig.~\ref{fig:hps_trigger_cal}):
 \begin{itemize}
 \item FADC (pulse finding): samples and digitizes the signal pulses from each detector channel. Sends the measured pulse energy and arrival time to the CTP.
\item CTP (cluster finding): groups FADC pulses from each half of the ECal into clusters. The cluster energy, arrival time, and hit pattern are sent to the SSP.
 \item SSP (cluster pair finding): searches for time coincidence of pairs of clusters from the top and bottom half of the ECal and applies topological selections.
 \end{itemize}
 \begin{figure}[t]
\begin{center}
{\small
 \includegraphics[width=7cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes the detector channel signals and sends them for cluster finding in the CTP. The CTP clusters are sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to readout the sub-detectors.}
 \label{fig:hps_trigger_cal}
}
\end{center}
 \end{figure}
The time coincidence window of pairs of clusters in the top and bottom half of the ECal are programmable 
with 4~ns resolution.

As described above in Sec.~\ref{sec:fadc}, the first stage of the trigger logic is incorporated into the 
FPGA's on the FADC boards which sends crystal energy and time information to the CTP. With the available 
3-bit time information, the CTP looks for time coincidence of crystal signals with 8~ns resolution (limit is 4~ns).
The cluster finding algorithm is very fast and makes use of the parallel processing nature of FPGA's by 
simultaneously searching for 125 clusters, up to 3x3 in size, across the calorimeter crystal array and performs 
the following tasks:
\begin{itemize}
\item Adds energy from hits together for every 3x3 square of channels in ECal.
\item Hits are added together if they occur (leading edge) within a programmable number of 4~ns clock cycles of each other (HPS will use 8~ns time coincidence time interval).
\item If the 3x3 energy sum is larger than the programmable cluster energy threshold and the sum is greater than any neighboring 3x3 windows, the CTP reports the cluster parameters to the SSP. 
\end{itemize}
The CTP evaluates all hits in its half of the calorimeter every 4~ns. A programmable time window is used to 
allow hits that are out of time with each other to be considered as part of a cluster sum. This is done by 
reporting hits when they occur and then reporting them again for the next $N$ number of 4~ns clock cycles, 
where $N \in [0,7]$. This is useful to deal with skew and jitter that develop from the detector, cabling, and 
electronics. As described above, the CTP only selects the 3x3 window with the highest energy sum of its 
neighbors. This filtering is applied to deal with overlapping clusters and cases where the cluster is larger than 
a 3x3 window.

 The final trigger decision is made in the CTP and Sub-System Processor (SSP). The Trigger Supervisor generates all necessary signals and controls the entire DAQ 
system readout through the Trigger Interface (TI) units. The TI units are installed in every crate that participate 
in the readout process.

@@

>>>>

\subsection{SVT Digitization, Event Building and Data Transmission}
\label{sec:svt_daq}
The SVT data acquisition (DAQ) based on the Reconfigurable Cluster Element (RCE) and Cluster 
Interconnect (CI) concept developed at SLAC as generic building blocks for DAQ systems in high energy 
physics {\bf [need reference]}. 
The RCE is a generic computational building block, housed on a separate daughter card called 
Data Processing Module (DPM), that are realized on an ATCA front board called the Cluster On Board (COB), 
see Fig.~\ref{fig:cob}.
 \begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/svt_daq_module}
	\caption{Picture of the SVT DAQ COB board with four data processing daughter cards visible.}
	\label{fig:cob}
}
\end{center}
\end{figure}
The Gen-I RCE used in the Test run consisted of a Virtex~5 FPGA with 1~GB of DDR3 RAM. 
The outputs of 12 SVT half-modules 
are digitized on the Rear-Transition-Module (RTM), a custom board on the back side of the ATCA crate, 
interfacing the HPS readout to the generic DAQ components on the COB. 
A pre-amplifier converts the APV25 differential current output to a different voltage output scaled to the sensitive 
range of a 14-bit ADC operating at the system clock of 41.667~MHz. 
The RTM is organized into four sections with each section supporting three SVT half-module hybrids (15 
APV25 channels). The RTM also includes a 4-channel Fiber Optic module and supporting logic which can be 
used to interface to the JLab trigger system supervisor.
Each section of the RTM, up to three hybrids, is input to a DPM which apply thresholds for data reduction and 
organizes the sample data into UDP datagrams. The DPM also hosts an I2C controller used to configure and 
monitor the APV25 chips on those three hybrids. A single ATCA crate with two COB cards was used, one 
supporting four DPMs and one supporting 3 DPMs and one DPM that is configured as the trigger and data 
transmission module. The two COB cards and their DPMs are interconnected with a 10Gb/s switch 
card~\cite{Larsen:2011zb} which 
also hosts two 1Gb/s Ethernet interfaces to the external SVT DAQ PC.

@@

After a trigger is received, six samples of the corresponding output of the APV25 shaper circuit are digitized. 
The samples from every channel on a sensor surviving the 
data reduction algorithm (see Sec.~\ref{sec:svt_daq}) are the basis for offline hit reconstruction. 
The six samples of the APV25 shaper output 
are fitted to an ideal $CR-RC$ function to extract the amplitude and the $t_0$ of the hit. 
The typical pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape} also demonstrates that the SVT 
was well timed in to the trigger with the rise of the pulse at the 3rd sampling point.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/08062012_run1351_samples_vs_amplitude}
    	\caption{The six pedestal subtracted samples from individual channels associated with a hit on a 
	track.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}

@@

\section{Electromagnetic Calorimeter}
\label{sec:ecal}

The electromagnetic calorimeter, installed downstream of the tracker, performs two essential functions for the experiment: provides a trigger to DAQ system and 
is used in the analysis for electron identification. 
%The device is highly segmented. It is fast, able to readout at rates comparable to those in the tracker, and able to provide good spatial and energy information to the trigger electronics. Like the tracker, the electromagnetic calorimeter is split to avoid impinging on the “dead zone”. The beam and radiative secondaries pass between the two of the calorimeter in vacuum, to avoid generating unnecessary backgrounds.
%Step\subsection{Purpose and Design Requirements}
%StepThe electromagnetic calorimeter (ECal) provide the trigger for the data acquisition and will be used for electron identification during the data analysis. The ECal is positioned after the analyzing dipole magnet and cover the full acceptance region of the SVT. It consist of two halves, split in the horizontal beam plane. The gap between upper and lower parts, $XXXX$~mrad (15mrad for electron running),  as seen from the target location, is necessary to avoid the "dead zone" where occupancy and radiation prevents any instrumentation to be situated.  
The energy of electrons of interest in HPS experiment will be in the range 
$0.5-6.5$~GeV. The ECal modules must have sufficient radiation lengths to absorb the full energy of these 
scattered electrons, have fine enough granularity to handle a high rate of electromagnetic 
background and survive large doses of radiation, especially near the beam plane. In addition, it needs to 
have excellent hit time resolution, $<10$~ns, in order to beat down beam backgrounds.

Requirements of a compact design, operation at high radiation and high rate environment, in the presence of a 
high magnetic field are fulfilled by a lead-tungstate (PbWO$_{4}$) crystal calorimeter with magnetic field 
resistant Avalanche Photodiode (APD) readout.  Such calorimeter has been operational in the inner 
part of the CLAS detector for experiments in Hall-B at JLab since 2005 and meet all requirements set by 
HPS. The modules from this calorimeter have been subsequently repurposed for the ECal.

@@

@@

%Step\subsubsection{PbWO$_4$ Crystals}
The HPS test run ECal uses PbWO$_{4}$ crystals, APDs, and preamplifier boards from the inner calorimeter (IC) of the CLAS detector~\cite{2008erhm.conf..421N}.  In Fig.~\ref{fig:ecal-module} a schematic view of the module assembly is shown. Lead-tungstate crystals are $160$ mm long and are tapered with front face of $13.3\times13.3$~mm$^2$, tapered to $16\times16$~mm$^2$ at the rear face. The crystals were fabricated with very tight tolerances: $19~\mu$m for 13.3~mm dimension and $13~\mu$m for the 16~mm dimension. Each crystal is wrapped in VM2000 multilayer polymer mirror film.
The APDs (Hamamatsu S8664-55) with $5\times5$~mm$^2$ active area and $75\%$ quantum efficiency were glued on the rear face of the crystals using MeltMount 1.7 thermal plastic adhesive. The low gain of APDs ($\sim 200$) was compensated with custom made preamplifier boards, which provide factor 2000 amplification of the ADP signal.

\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}

\subsection{Layout}

Like the tracker, the electromagnetic calorimeter is split into two parts, beam-up and beam-down, to avoid impinging on the “dead zone”. The two parts of the Ecal are separated by a vacuum chamber, see Fig.~\ref{fig:ecal}. The beam and radiative secondaries pass between the two parts in vacuum, to avoid generating unnecessary backgrounds. The crystal modules in each part are arranged in rectangular formation, as shown in top panel of Fig.~\ref{fig:ecal}. There are 5 layers in each formation. Each layer has $46$ crystals except for the layers closest to the beam 
plane where 9 modules were removed to allow a larger opening for the outgoing electron and photon beams. 
In order to have stable light yield from crystals, and stable operation of APDs and preamplifiers, the module assembly of each ECal part was enclosed inside a temperature controlled box.

Inside the box PbWO$_{4}$ crystals were supported by aluminum frames stacked on top of each other.  The bias voltage for APDs ($\sim 400$ V), the low voltage for preamplifiers ($\pm 5$ V), and the signals from APDs were transmitted in and out of the enclosure through a backplane called motherboard. The $221$ modules in each part of the ECal were divided into two low voltage groups and 12 bias voltage groups. Modules in bias voltage groups were selected to have gain uniformity of $\sim 20\%$.

During the test run ECal was mounted downstream of the analyzing dipole magnet at the distance of about $147$~cm from the 
upstream edge of the magnet. 
%The two ECal modules are positioned just above and below the ECal vacuum chamber as shown in bottom panel of Fig.~\ref{fig:ecal}. 
%The primary electron beam, scattered electrons and bremsstrahlung photons pass through the system in vacuum. 
The gap between innermost edge of the crystals of two parts was just $7.4$~cm from the beam. The temperature stability of the system during the test run was better than $1^\circ~$F, the temperature uniformity inside the box was better than  $4^\circ~$F.

\begin{center}
{\small
\begin{figure}
\includegraphics[width=7cm]{figures/ECal}
\includegraphics[width=7cm]{figures/ecal_vacuumchamber.pdf}
%\includegraphics[width=7cm]{figures/ecal_assembly}
\caption{Layout of PbWO$_{4}$ modules in two halves of the ECal (top) and ECal assembly together with ECal vacuum chamber (bottom). }
\label{fig:ecal}
\end{figure}
}
\end{center}
%Step \subsubsection{Support Structure}
%Step In order to stabilize the calorimeter's performance, the crystals, APDs, and amplifiers are enclosed within a temperature controlled environment, held constant at the level of .
%Step\subsubsection{Avalanche Photodiode Readout and Preamplifiers}
%Step\label{sec:apd}
%StepAvalanche Photodiodes (APD) S8664-55 produced by Hamamatsu Corp. are used as photodetectors. They have a $5\times5$~mm$^2$ active area and quantum efficiency of about 75\%. Single APDs were centered on the back ends of the crystals using MeltMount 1.7 thermal plastic adhesive. The low? amplitude signals of APDs are amplified using custom?made preamplifier boards. 
%The gain uniformity of the APDs is on the order of XXXX\%; with APD high voltage bias varying by as much as 100V. Since only a limited number of high voltage channels were available during the test, 10 channels were grouped together with an attempt to group modules with similar gains.

\subsection{Signal readout}
\label{sec:fadc}
Amplified APD signal from the motherboard is sent to a $2:1$ signal splitter. After the split, 2/3 of the signal is sent a discriminator then to a Time-to-Digital 
Converter (TDC). The 1/3 of the signal is fed to a single channel of 16-channel JLAB Flash Amplitude-to-Digital Converter VXS module (FADC250)~\cite{fadc}, see Fig.\ref{fig:fadc}.  
\begin{figure}[t]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/FADC250_Photo_001.jpg}
\caption{A Jefferson Lab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
Two 20-slot VXS crates with 14 FADC bards were needed to accommodate the system: one for each half of the ECal with 221 
channels.

The FADCs store 12-bit digitized samples at 250~MHz in 8~$\mu$s deep pipelines. 
When a trigger is received, the appropriate part of the pipeline is accessed for the readout. HPS used FADCs in the integration mode. A predefined number of samples of the signal that exceeds a predefined threshold have been summed and stored in 17-bit register for readout. The number of samples for integration was defined by two parameters that could be set separately, the number of samples integrated before the signal crossed threshold ($NSB$) and  the number of samples integrated after the signal crossed threshold ($NSA$), see Fig.\ref{fig:hps_trigger_data}. 
This scheme significantly compresses the data input 
to the FADC. During data analysis, a pedestal value is subtracted to obtain the actual summed energy.

\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/hps_trigger_data}
\caption{\small{FADC data paths}}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}

subsection{Calibration and Performance}
\label{sec:ecal_calibration}

For various reasons only 385 out of 442 modules (87\%) were operational during the test run.  Of 442 modules, 39 were disabled or disconnected and were not read out by the DAQ. 
13 of these were not read out because of a shortage of FADC readout boards (see Sec.~\ref{sec:fadc}).
The remainder either had no HV bias on the APD, or were disabled in the FADC software due to noise.
In the data, we identified two types of abnormal channels. 
One FADC was not sending trigger signals correctly, resulting in low efficiency.
 This affected the 13 channels read out by that FADC. 5 channels were diagnosed as noisy because they 
 had a high incidence of hits out of coincidence with the trigger. A large number of channels were originally 
 misidentified as noisy because they had much higher hit occupancy than neighboring channels.
Gain calibration shows that these channels have high gain (and thus lower energy threshold) but are otherwise normal. The abnormal channels were ignored in analysis in order to simplify comparison with Monte Carlo. This leaves 385 useful channels---87\% of the ECal.

@@

We calibrate gain of the individual ECal channels using the SVT measurement of track momentum and 
comparison to Monte Carlo simulation.  We disable all SVT and ECal channels in the simulation that were 
inoperable or noisy in the test run, so any efficiency or bias effects that affect the real data should be reflected 
in the simulation as well; then we use a formula to compute the ``weighted E/p'' for a crystal, representing the average E/p for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$
, and iteratively adjust the gains until the weighted E/p is equal for test run data and simulation.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/eoverp_simulation}
	\caption{Weighted E/p from Monte Carlo simulation after gain calibration.}
	\label{fig:gains}
}
\end{center}
\end{figure}

@@

%Step\subsection{Monitoring}

@@

\section{Trigger and Data Acquisition}

The HPS Test apparatus DAQ handles the acquisition of data from two sub-detectors: the 
SVT and the ECal, with two DAQ architectures. The SVT is readout with Advanced 
Telecom Communications Architecture (ATCA) hardware while the ECal uses VXS based 
hardware. The trigger system receives input from the ECal and distributes a trigger signal 
to all detector sub- systems to read out a selected event.

\subsection{Trigger system}
\label{sec:trigger}

The HPS test run trigger system is designed to efficiently select \ee{} events by 
using information from the ECal . The trigger looks for time coincidences of clusters in the top and bottom half 
of the ECal.  The trigger system can be broken down into three sections (see Fig.~\ref{fig:hps_trigger_cal}):
 \begin{itemize}
 \item FADC (pulse finding): samples and digitizes the signal pulses from each detector channel. Sends the measured pulse energy and arrival time to the Crate Trigger Processor (CTP),
\item CTP (cluster finding): groups FADC pulses from each half of the ECal into clusters. The cluster energy and arrival time are sent to the Sub-System Processor (SSP),
 \item SSP (cluster pair finding): searches for time coincidence of pairs of clusters from the top and bottom half of the ECal and applies topological selections.
 \end{itemize}
 \begin{figure}[t]
\begin{center}
{\small
 \includegraphics[width=7cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes the detector channel signals and sends them for cluster finding in the CTP. The CTP clusters are sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to readout the sub-detectors.}
 \label{fig:hps_trigger_cal}
}
\end{center}
 \end{figure}
The time coincidence window of pairs of clusters in the top and bottom half of the ECal are programmable 
with 4~ns resolution.

As described above, the first stage of the trigger logic is incorporated into the 
FPGA's on the FADC boards. 
For the trigger path, every channel has, in addition to $NSB$ and $NSA$: 
\begin{itemize}
% \item number of samples integrated before the threshold crossing (NSB),
 %\item number of samples integrated after the  threshold crossing (NSA),
 \item trigger threshold, measured in ADC counts, 
 \item a pedestal, 
 \item a conversion factor (gain) that converts the ADC counts to energy in MeV (with 13 bits:  from 0 to 8191~MeV), 
 \item an energy discriminator threshold (minimum energy cutoff).
 \end{itemize}
Note that the threshold for the trigger path can be set independently from the readout threshold.
The pedestal value is subtracted from the integrated sum over $NSB+NSA$ samples and 
converted to MeV units using a supplied gain conversion factor. The energy discriminator can be used to cut 
off low energy pulses before reporting to CTP. 
The values reported to the CTP are the 5-bit pulse energy and 3-bit time information at which the 
pulse crossed the threshold. Data for every channel is sent to CTP every 32~ns. With the available 
3-bit time information, the CTP looks for time coincidence of crystal signals with 8~ns resolution (limit is 4~ns).
The cluster finding algorithm is fast and makes use of the parallel processing nature of FPGA's by 
simultaneously searching for 125 clusters, up to 3x3 in size, across the calorimeter crystal array and performs 
the following tasks:
\begin{itemize}
\item Adds energy from hits together for every 3x3 square of channels in ECal.
\item Hits are added together if they occur (leading edge) within a programmable number of 4~ns clock cycles of each other (HPS test run used 8~ns coincidence time interval).
\item If the 3x3 energy sum is larger than the programmable cluster energy threshold and the sum is greater than any neighboring 3x3 windows, the CTP reports the cluster parameters to the SSP. 
\end{itemize}
The CTP evaluates all hits in its half of the calorimeter every 4~ns. A programmable time window is used to 
allow hits that are out of time with each other to be considered as part of a cluster sum. This is done by 
reporting hits when they occur and then reporting them again for the next $N$ number of 4~ns clock cycles, 
where $N \in [0,7]$. This is useful to deal with skew and jitter that develop from the detector, cabling, and 
electronics. As described above, the CTP only selects the 3x3 window with the highest energy sum of its 
neighbors. This filtering is applied to deal with overlapping clusters and cases where the cluster is larger than 
a 3x3 window.

 The final trigger decision is made by CTPs and SSP passed to the Trigger Supervisor (TS). The TS generates all necessary signals and controls the entire DAQ 
system readout through the Trigger Interface (TI) units. The TI units are installed in every crate that participate 
in the readout process.

@@

During the test run, for the most part trigger system run in single cluster mode and was tested to work for up to 20 kHz trigger rates.

\subsection{Data Acquisition and Online Computing}
\label{sec:daq}
For the ECal, every VXS crate contains a Readout Controller (ROC)
that collects digitized information, processes it, and sends it on to the Event Builder (EB). The ROC is a single 
blade Intel-based CPU module running DAQ software under CentOS Linux OS. For the SVT ATCA system, 
the ROC application runs on an embedded processor situated on the ATCA main board. The EB assembles 
information from the SVT and ECal ROCs into a single event which is passed to the Event 
Recorder (ER) that writes it to a RAID5-based data storage system capable of handling up to 100 MB/s. The 
EB and other critical components run on multicore Intel-based multi-CPU servers. The DAQ network system is 
a Foundry router providing high-speed connections between the DAQ components and to the JLab 
computing facility. The SVT ROC, which must handle large data volumes, has a 10 Gbit/s link to the Foundry 
router, while a 1 Gbit/s link is adequate for the ECal. A 10 Gbit/s uplink to the JLab 
computing facility is used for long-term storage.
